Towards atomistic resolution structure of phosphatidylcholine glycerol backbone and choline headgroup at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-03-31 HGmodelMANUSCRIPT/HGmodel_ACStemplate.tex
update hydration results in the manuscript

\begin{table}[htb]
\centering
\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ Water/lipid molar ratio
$^b$ The number of lipid molecules
$^c$ The number of water molecules
$^d$ Simulation temperature
$^e$ The total simulation time
$^f$ Time frames used in the analysis
$^g$ Reference link for the downloadable simulation files
$^h$ Reference for the full simulation details
}\label{systemsDEHYD}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
CHARMM36\cite{klauda10}              & POPC   & 31 & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
MacRog\cite{maciejewski14}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
\end{tabular}
\end{table}

@@

\subsubsection{Phospholipid bilayer with low hydration level}
The experimental order parameters available in the literature~\cite{dvinskikh05b,ulrich94,bechinger91} 
for the glycerol backbone and choline as a function of hydration level are shown in Fig.~\ref{ordPhydr}. 
The independently reported values for choline segments are in good agreement with each other (despite 
slight differences in temperature and acyl chain composition),
showing increase for both segments with decreasing hydration level. It should be noted that only 
absolute values were measured in the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91}, but
we have included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
Consequently, the $\beta$ order parameter with negative sign actually increases with dehydration 
since the absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
Slight decrease for the glycerol backbone order parameters were observed with dehydration~\cite{dvinskikh05b}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf}
  \todo{It should be checked that we have the up-to-date GAFF results in this plot.} \\
  \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
    Note that to elucidate the relative change as a function of hydration level,
    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
    matches (within $\pm$0.01) the experimental value at full hydration;
    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
  }
\end{figure}

@@

It should be noted, however, that free energy landscape is not correct in the models which
are not able to reproduce the experimental order parameters. Thus the energetical response
to the dehydration may not be correct even though the order parameter response is qualitatively correct.
This issue may have some influence on the dehydration energetic calculations made with the Berger model~\cite{eun09,schneck12}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydrationPN.pdf}

>>>>

\begin{table}[htb]
\centering
\caption{Simulated single component lipid bilayers with varying hydration levels. The simulation file data sets marked with $^*$ include also part of the trajectory.
$^a$ Water/lipid molar ratio
$^b$ The number of lipid molecules
$^c$ The number of water molecules
$^d$ Simulation temperature
$^e$ The total simulation time
$^f$ Time frames used in the analysis
$^g$ Reference link for the downloadable simulation files
$^h$ Reference for the full simulation details
}\label{systemsDEHYD}
\begin{tabular}{c c c c c c c c c c}
%\hline
Force field & lipid & $^a$n (w/l)   & $^b$N$_{\rm l}$   &  $^c$N$_{\rm w}$ & $^d$T (K)  & $^e$t$_{{\rm sim}}$(ns)  & $^f$t$_{{\rm anal}}$ (ns)& $^g$Files  &  $^h$Details\\
\hline
Berger-POPC-07~\cite{ollila07a}          &   POPC & 57  &128 & 7290  & 298  & 270 & 240 & [\citenum{bergerFILESpopc}]$^*$ & SI \\
                                        &   POPC & 7  &128 & 896   & 298  & 60 & 50 & [\citenum{bergerDEHYDfiles}]$^*$ & SI \\
CHARMM36\cite{klauda10}          & POPC   & 40 & 128 &  5120   & 303 & 150 & 100  & [\citenum{charmm36files}]$^*$   & SI \\
                                 & POPC   & 31  & 72  &  2242 & 303 & 30 & 20 & [\citenum{charmm36filesSHORT}]$^*$ & SI \\
                               & POPC   & 15 & 72 &  1080  & 303 & 59 & 40 & [\citenum{charmm36files15wPERl}]$^*$ & SI \\
                            & POPC   & 7  & 72  &  504  & 303 & 60 & 20 & [\citenum{charmm36files7wPERl}]$^*$ & SI \\
MacRog\cite{maciejewski14}     & POPC   & 50 & 288  & 14400 & 310 & 90 & 40 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                               & POPC   & 25 & 288  & 7200 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 20 & 288  & 5760 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 15 & 288  & 4320 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 10 & 288  & 2880 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
                                & POPC   & 5 & 288   & 1440 & 310 & 100 & 50 & [\citenum{macrogdehydFILES}]$^*$ & SI \\
GAFFlipid\cite{dickson12}      & POPC   & 31& 126  & 3948  & 303 & 137 & 32 & [\citenum{GAFFlipidFILES}]$^*$ & SI  \\
                               & POPC   & 7& 126  & 896   & 303 & 130 & 40 & [\citenum{gaffDEHYDfiles}]$^*$ & SI \\
\end{tabular}
\end{table}

@@

\subsubsection{Phospholipid bilayer with low hydration level}
The experimental order parameters available in the literature~\cite{dvinskikh05b,ulrich94,bechinger91} 
for the glycerol backbone and choline as a function of hydration level are shown in Fig.~\ref{ordPhydr}. 
The independently reported values for choline segments are in good agreement with each other (despite 
slight differences in temperature and acyl chain composition),
showing increase for both segments with decreasing hydration level. It should be noted that only 
absolute values were measured in the original experiments~\cite{dvinskikh05b,ulrich94,bechinger91}, but
we have included the signs measured in separate studies~\cite{hong95a,hong95b,gross97}. 
Consequently, the $\beta$ order parameter with negative sign actually increases with dehydration 
since the absolute value decreases~\cite{dvinskikh05b,ulrich94,bechinger91}.
Slight decrease for the glycerol backbone order parameters were observed with dehydration~\cite{dvinskikh05b}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../DATAreportediINblog/dehydration.pdf} \\
  \todo{DONE} \\
  \caption{\label{ordPhydr}
    The effect of dehydration on glycerol and choline order parameters in experiments.
    The magnitudes of order parameters are measured for DMPC ($^{13}$C NMR) at 314~K~\cite{dvinskikh05b}, 
    for POPC ($^2$H NMR) at 296~K~\cite{bechinger91} and for DOPC ($^2$H NMR) at 303~K~\cite{ulrich94}. 
    The signs are based on the measurements by Hong et al.~\cite{hong95a,hong95b} 
    and Gross et al.~\cite{gross97}.
    Note that to elucidate the relative change as a function of hydration level,
    the simulation results are shifted such that the (smaller) $S_\mathrm{CH}$
    matches (within $\pm$0.01) the experimental value at full hydration;
    the shift magnitudes for each of the force fields are listed ($S_\mathrm{CH}$+shift) in the y-label.
  }
\end{figure}

@@

It should be noted, however, that free energy landscape is not correct in the models which
are not able to reproduce the experimental order parameters. Thus the energetical response
to the dehydration may not be correct even though the order parameter response is qualitatively correct.
This issue may have some influence on the dehydration energetic calculations made with the Berger model~\cite{eun09,schneck12}.
\begin{figure}[]
  \centering
  \includegraphics[width=8.6cm]{../Fig/dehydrationPN.pdf}
